### 2016-03-03 Draft.tex
Routine additions.

However, the assumption of a constant detection rate throughout the observation period is not always justified.  Animal behaviors such as bout singing or intermittent singing in birds and frogs or diving in whales can lead to time-varying rates of detection \citep{Scott2005, Diefenbach2007, Reidy2011}.  The act of observation may affect animal behavior.  Depending upon the study species, the presence of an observer may temporarily either stimulate or suppress detectable cues from the studied species \citep{McSheaRappole1997, Rosenstock2002}.  Survey methodology itself can affect detection rates.  In the presence of detection heterogeneity, easily detected animals will be removed first, leaving only harder to detect animals, thus resulting in a marginal detection rate that varies over time.  Observers may also become distracted as the survey progresses \citep{Johnson2008}.  And, depending upon protocol, observers may become aware of animals during their settling-in period, resulting in elevated counts early during in the survey period \citep{LeeMarsden2008}.

@@

The framing of our model in terms of time-to-detection leads to two practical differences vis-a-vis constant-detection models.  First, in order to model covariate and random effects on detection, we perform mixed effects linear regression on the log of the rate parameter as in \citet{Solymos2013}.  Most existing studies have instead constructed regression models on the logit of the equal-interval detection probability \adam{Amundson2014}.  Such an approach is not possible when detection rates are not constant.  \adam{This statement is false.  It has been done, but the form of the non-constancy must be proscribed.  See Schnute 1983, Scruton and Gibson (ref: Mantyniemi), and Mantyniemi2015 for examples}.  Second, because we can obtain interval-specific detection probabilities from the TTDD by partitioning its cdf, we can directly model the data according to their existing interval structure rather than subdividing the observation period into intervals of equal duration.  Indeed, with a few simplifications, our model fits exact time-to-detection data, whereas existing constant-detection removal models only approximate exact data by subdividing the observation interval into a large number of fine intervals \citep{Reidy2011, Amundson2014}.

@@

The utility of applying a mixture TTDD for constant-detection models has been identified several times \citep{Pledger2000, Farnsworth2002, Alldredge2007, Etterson2009, Reidy2011}\adam{Alldredge, Farnsworth, Pledger, Reidy verified}.  The mixture allows an analysis to account for detection heterogeneity across easy- and hard-to-detect subgroups of a population.  Even so, some recent studies have assumed the absence of a mixture \citep{Solymos2013, Amundson2014}.  Our study indicates that mixtures are useful even when a non-constant TTDD is used for hard-to-detect individuals.  Models that include a mixture generally perform well even when no heterogeneity is present, but models that omit a mixture are biased and overly certain when detection heterogeneity exists.  Indeed, for peaked datasets, mixture models can outperform non-mixture models even in the absence of heterogeneity.

>>>>

However, the assumption of a constant detection rate throughout the observation period is not always justified \citep{Alldredge2007}.  Animal behaviors such as bout singing or intermittent singing in birds and frogs or diving in whales can lead to time-varying rates of detection \citep{Scott2005, Diefenbach2007, Reidy2011}.  The act of observation may affect animal behavior.  Depending upon the study species, the presence of an observer may temporarily either stimulate or suppress detectable cues from the studied species \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007}.  Survey methodology itself can affect detection rates.  In the presence of detection heterogeneity, easily detected animals will be removed first, leaving only harder to detect animals, thus resulting in a marginal detection rate that varies over time \citep{Farnsworth2005}.  Observers may also become distracted as the survey progresses \citep{Johnson2008}.  And, depending upon protocol, observers may become aware of animals during their settling-in period, resulting in elevated counts early during in the survey period \citep{LeeMarsden2008}.

@@

The framing of our model in terms of time-to-detection leads to two practical differences vis-a-vis constant-detection models.  First, in order to model covariate and random effects on detection, we perform mixed effects linear regression on the log of the rate parameter as in \citet{Solymos2013}.  Most existing studies have instead constructed regression models on the logit of the equal-interval detection probability \adam{Amundson2014}.  Such an approach is not possible when detection rates are not constant.  Second, because we can obtain interval-specific detection probabilities from the TTDD by partitioning its cdf, we can directly model the data according to their existing interval structure rather than subdividing the observation period into intervals of equal duration.  Indeed, with a few simplifications, our model fits exact time-to-detection data, whereas existing constant-detection removal models only approximate exact data by subdividing the observation interval into a large number of fine intervals \citep{Reidy2011, Amundson2014}.

\adam{Based on further lit review, I think I need to insert a paragraph here about what time-varying modeling has been done.  In the fisheries world, models of time-varying catchability have been implemented -- see Schnute 1983, Scruton and Gibson (ref: Mantyniemi), and Mantyniemi2015 for examples.  Also, Alldredge 2007 implements a complete-history of detection model (technically mark-recapture, not removal) that estimates interval-specific detection probabilities (complete with covariates and detection heterogeneity).  However, it cannot be applied in a removal context, because removal data contains only first times to detection, and the Alldredge model requires subsequent observations -- removal models compensate by specifying a TTDD, thus enabling us to extrapolate beyond the observation period.}

@@

The utility of applying a mixture TTDD for constant-detection models has been identified several times \citep{Pledger2000, Farnsworth2002, Alldredge2007, Reidy2011}.  The mixture allows an analysis to account for detection heterogeneity across easy- and hard-to-detect subgroups of a population.  Even so, some recent studies have assumed the absence of a mixture \citep{Solymos2013, Amundson2014}.  Our study indicates that mixtures are useful even when a non-constant TTDD is used for hard-to-detect individuals.  Models that include a mixture generally perform well even when no heterogeneity is present, but models that omit a mixture are biased and overly certain when detection heterogeneity exists.  Indeed, for peaked datasets, mixture models can outperform non-mixture models even in the absence of heterogeneity.
